---
title: "Simple Analyzer"
author: "Mohamed Tleis"
date: "7/14/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Simple Analyzer : Overview
- A simple data Analyzer. A simple csv file can be uploaded where the first two columns are analyzed. The first column holds the response and the second column contains the main predictor feature.

```{r simpleAnalyzer}
  
```

## Input & Data Viewer
- One the side panel, there is an Input tab, where a file can be uploaded.
- The contents of the first two columns of the input file can be viewed in the Data tab in the main panel.
- When a file is uploaded the histogram and the scatter plot are reactively updated.
- When no file is present. A sample dataset from the default 'trees' data is loaded. The response column is the Volume feature. Main predictor is the Girth feature.
```{r, eval = TRUE}
  data(trees)
  data.frame(trees$Volume, trees$Girth)
```


## Histogram
- In the main panel, Histogram tab show a histogram of the main predictor.
- The number of bins can be controlled via the Slider (located under Slider tab in the side panel).
```{r eval=TRUE}
  
```

## Scatter Plot
- In 'lm-Plot' tab, the response is plotted on the y-axis and the predictor on the x-axis.

## Regression Model
- Whenever a brushed dataset is selected, a lm line is plotted on the chart.
