---
title: "Simple Analyzer"
author: "Mohamed Tleis"
date: "7/14/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
```

## Simple Analyzer : Overview
- A simple data Analyzer. 
- A simple csv file can be uploaded where the first two columns are analyzed.
- The first column holds the response and the second column contains the main predictor feature.
- Access App at : 

```{r simpleAnalyzer, echo = TRUE}
   https://mtleis.shinyapps.io/simple_analyzer/
```

## Input & Data Viewer {.smaller}
- One the side panel, there is an Input tab, where a file can be uploaded.
- The contents of the first two columns of the input file can be viewed in the Data tab in the main panel.
- When a file is uploaded the histogram and the scatter plot are reactively updated.
- When no file is present. A sample dataset from the default 'trees' data is loaded. The response column is the Volume feature. Main predictor is the Girth feature.

```{r, eval = TRUE , echo = FALSE}
  data(trees)
  head(data.frame(trees$Volume, trees$Girth))
```


## Histogram {.smaller}
- In the main panel, Histogram tab shows a histogram of the response feature.
- The number of bins can be controlled via the Slider (located under Slider tab in the side panel).

```{r histogram, eval=TRUE, echo = FALSE, fig.width=6, fig.height=3}
   hist(trees$Volume, breaks = 20, col = 'darkgray', border = 'white',
        main="Response Feature")
```

## Scatter Plot
- In 'lm-Plot' tab, the response is plotted on the y-axis and the predictor on the x-axis.
```{r scatter, eval=TRUE, fig.width=6, fig.height=3}
  plot(trees$Girth, trees$Volume, xlab="Main Predictor", ylab="Response",
       main="Predictor", cex=1.5, pch=16, bty="n")
```

## Regression Model
- Whenever a brushed dataset is selected, a lm line is plotted on the chart.

```{r lm, eval=TRUE, fig.width=6, fig.height=3}
  plot(trees$Girth, trees$Volume, xlab="Main Predictor", ylab="Response",
       main="Predictor", cex=1.5, pch=16, bty="n")

  abline(lm(trees$Volume~trees$Girth, data=trees), col='red', lwd=2)
```
